---
title: "Synaptic_propagation_simulation"
output: html_document
date: "2024-11-08"
---
# load libraries
```{r message=FALSE, warning=FALSE}
library(psych)
library(reticulate)
library(corrplot)
library(lm.beta)
library(boot)
library(plyr)
library(tibble)
library(lme4)
library(lmerTest)
library(NbClust)
library(zoo)
library(ggplot2)
library(ggsci)
library(Hmisc)


# set virtual environment 
reticulate::use_python("/Users/yluan/.virtualenvs/r_virtualenv/bin/python3.8/", required = TRUE) ## set the path to python under virtual environment 

# load toolbox for generating surrogate maps for spin test
# reticulate::py_install("netneurotools", pip_options="--no-binary='netneurotools'", ignore_installed=TRUE)
# reticulate::py_install("neuromaps", pip_options="--no-binary='neuromaps'", ignore_installed=TRUE)
np  <- import("numpy", convert = F)
nnt <- import("netneurotools")
nntdata <- import("netneurotools")$datasets

nm <- import("neuromaps")
nmnulls <- import("neuromaps")$nulls
nmimgs <- import("neuromaps")$images
nmparc <- import("neuromaps")$parcellate

# convert the schaefer atlas to a gifti file
schaefer = nntdata$fetch_schaefer2018('fslr32k')['200Parcels7Networks']
parc <- nmimgs$dlabel_to_gifti(schaefer)



# lead Schaefer2018_200Parcels_7Networks_order_numeric.txt 
# set the path to schaefer200x7CommunityAffiliation_corrected.txt
sf.assign <- read.table("/Volumes/Local/Ying_L/Data_processing/atlas/Schaefer2018_200Parcels/schaefer200x7CommunityAffiliation_corrected.txt") 
colors.Schaefer.networks <- c("forestgreen" ,"indianred2" ,"darkgoldenrod1" ,"lemonchiffon1" ,"skyblue3" ,"mediumorchid1" ,"magenta4")[c(3,7,6,5,2,4,1)]

```



# ----- SIMULATE DATA ------

## define numbers

```{r}
number_of_ROIs = 200
number_of_subjects = 124
```

## simulate connectivity matrix
```{r}
# functional connectivity matrix
set.seed(123) 
fc.mat <- matrix(runif(number_of_ROIs * number_of_ROIs, min = -1, max = 1), nrow = number_of_ROIs, ncol = number_of_ROIs)
diag(fc.mat) <- 1
fc.mat <- fisherz(fc.mat)
fc.mat[which(fc.mat==Inf,arr.ind = T)] <- NA
corrplot(fc.mat, diag = FALSE, tl.pos = "n", tl.cex = 0.5, method = "color",is.corr = F)

# 1000 shuffled functional connectivity matrices
shuffled.fc <- lapply(1:1000, function(x){
  fc.mat.tmp <- matrix(runif(number_of_ROIs * number_of_ROIs, min = -1, max = 1), nrow = number_of_ROIs, ncol = number_of_ROIs)
  diag(fc.mat.tmp) <- 1
  fc.mat.tmp <- fisherz(fc.mat.tmp)
  fc.mat.tmp[which(fc.mat.tmp==Inf,arr.ind = T)] <- NA
  fc.mat.tmp
})

# Euclidean distance matrix
eu.d <- matrix(runif(number_of_ROIs * number_of_ROIs, min = 0, max = 15), nrow = number_of_ROIs, ncol = number_of_ROIs)
diag(eu.d) <- NA
```


## simulate subject characteristics and plasma ptau biomarker data
```{r}
set.seed(123) 

df <- data.frame(
    SubID = 1:number_of_subjects,
    Age.SDM8 = sample(60:90, number_of_subjects, replace = TRUE),
    SEX = sample(c("male", "female"), number_of_subjects, replace = TRUE),
    DX.AB = sample(c("CN_0", "CN_1","MCI_1", "Dementia_1"), number_of_subjects, replace = TRUE),
    pTau181 = runif(number_of_subjects),
    AV45.global.mean = runif(number_of_subjects, min = 0, max = 2.5)) 

df$DX <- mapply(function(x){strsplit(x,"_")[[1]][1]}, df$DX.AB)
df$AV45.bin <- mapply(function(x){as.numeric(strsplit(x,"_")[[1]][2])}, df$DX.AB)
```


## simulate SV2A-PET SUVRs and SV2A-PET w-score
```{r}
# Generate random SV2A PET SUVRs for each ROI for the same subjects
SV2A_PET_SUVRs <- matrix(runif(number_of_subjects * number_of_ROIs, min = 0, max = 2), nrow = number_of_subjects)
colnames(SV2A_PET_SUVRs) <- paste0("SDM8.V", 1:number_of_ROIs)

# Generate random SV2A PET w-score  for each ROI for the same subjects
SV2A_PET_wscore <- matrix(runif(number_of_subjects * number_of_ROIs, min = -2, max = 2), nrow = number_of_subjects)
colnames(SV2A_PET_wscore) <- paste0("W.SDM8.V", 1:number_of_ROIs)

# Generate random SV2A PET w-score change rates for each ROI for a subset of subjects
SV2A_PET_wscore_CR <- matrix(runif(25 * number_of_ROIs, min = -1, max = 0.5), nrow = 25)
colnames(SV2A_PET_wscore_CR) <- paste0("CR.W.SDM8.V", 1:number_of_ROIs)
SV2A_PET_wscore_CR <- as.data.frame(SV2A_PET_wscore_CR)
SV2A_PET_wscore_CR$SubID <- sample(1:124,25,replace = F)


# Combine the sv2a PET SUVRs and w score with the existing data frame
df <- cbind(df, SV2A_PET_SUVRs, SV2A_PET_wscore)
df <- merge(df, SV2A_PET_wscore_CR, by = "SubID", all.x = T)

# group-average SV2A w-score
df.ad <- subset(df, DX.AB %in% c("CN_1","MCI_1","Dementia_1"))
mean.w.ad <- colMeans(df.ad[,paste0("W.SDM8.V",1:200)])
mean.w.ad.surrogates <- nmnulls$vasa(mean.w.ad,atlas='fsLR',density='32k',
                                            parcellation=parc,seed=as.integer(45))

df.cn0 <- subset(df, DX.AB %in% c("CN_0"))
mean.w.cn0  <- colMeans(df.cn0[,paste0("W.SDM8.V",1:200)])
```




# ----- DATA  ANALYSES------

#---- Analysis 1: the association between sv2a covariance and functional connectivity matrix
```{r}

# generate the sv2s covariance matrix in abeta positive group
w.cov.ad <- cor(as.matrix(df.ad[,paste0("W.SDM8.V",1:200)]),method = "spearman")
w.cov.z.ad <- fisherz(w.cov.ad)
w.cov.z.ad[which(w.cov.z.ad==Inf)] <- NA

# generate the sv2s covariance matrix in cognitively normal abeta negative group
w.cov.cn0 <- cor(as.matrix(df.cn0[,paste0("W.SDM8.V",1:200)]),method = "spearman")
w.cov.z.cn0 <- fisherz(w.cov.cn0)
w.cov.z.cn0[which(w.cov.z.cn0==Inf)] <- NA


# prepare data fram for analyses
w.cov.vecterized.ad <- w.cov.z.ad[lower.tri(w.cov.z.ad)]
w.cov.vecterized.cn0 <- w.cov.z.cn0[lower.tri(w.cov.z.cn0)]
fc.verterized <- fc.mat[lower.tri(fc.mat)]
df.stat.cov <- data.frame(w.cov.ad= w.cov.vecterized.ad,
                          w.cov.cn0=w.cov.vecterized.cn0,
                          fc=fc.verterized)

## Abeta+ group
mod.lm.ad <- lm(data = df.stat.cov,w.cov.ad~fc)
mod.lm.ad <- summary(lm.beta(mod.lm.ad))
beta.val.ad <- mod.lm.ad$coefficients["fc","Standardized"]
mod.lm.ad

beta.shuffled.ad <- mapply(function(x){
  fc.mat.shuffled <- shuffled.fc[[x]]
  df.cov.shuffled <- data.frame(w.cov.ad= w.cov.vecterized.ad,
                                w.cov.cn0=w.cov.vecterized.cn0,
                                fc=fc.mat.shuffled[lower.tri(fc.mat.shuffled)])
  mod.lm.shuffled <- lm(data = df.cov.shuffled,w.cov.ad~fc)
  mod.lm.shuffled  <- summary(lm.beta(mod.lm.shuffled))
  beta.val.shuffled <- mod.lm.shuffled$coefficients[2,2]
}, 1:1000)
p.val.shuffled.ad <- length(which(beta.shuffled.ad > beta.val.ad))/length(beta.shuffled.ad)

## CN abeta- group
mod.lm.cn0 <- lm(data = df.stat.cov,w.cov.cn0~fc)
mod.lm.cn0 <- summary(lm.beta(mod.lm.cn0))
beta.val.cn0 <- mod.lm.cn0$coefficients["fc","Standardized"]
mod.lm.cn0

beta.shuffled.cn0 <- mapply(function(x){
  fc.mat.shuffled <- shuffled.fc[[x]]
  df.cov.shuffled <- data.frame(w.cov.ad= w.cov.vecterized.ad,
                                w.cov.cn0=w.cov.vecterized.cn0,
                                fc=fc.mat.shuffled[lower.tri(fc.mat.shuffled)])
  mod.lm.shuffled <- lm(data = df.cov.shuffled,w.cov.cn0~fc)
  mod.lm.shuffled  <- summary(lm.beta(mod.lm.shuffled))
  beta.val.shuffled <- mod.lm.shuffled$coefficients[2,2]
}, 1:1000)
p.val.shuffled.cn0 <- length(which(beta.shuffled.cn0 > beta.val.cn0))/length(beta.shuffled.cn0)


## comparison between Abeta+ group and CN abeta- group with bootstrapping
### define function
sv2a.cov.z.boot <- function(mat,ind){
  sv2a.mat <- as.matrix(mat[ind,paste0("W.SDM8.V",1:200)])
  w.cov.boot <- cor(sv2a.mat,method = "spearman")
  w.cov.z.boot <- fisherz(w.cov.boot)
  w.cov.z.boot[which(w.cov.z.boot==Inf)] <- NA
  
  w.cov.vecterized.boot<- w.cov.z.boot[lower.tri(w.cov.z.boot)]
  fc.verterized <- fc.mat[lower.tri(fc.mat)]
  
  df.cov.boot <- data.frame(w.cov.boot= w.cov.vecterized.boot,
                            fc=fc.verterized)

  mod.lm <- lm(data = df.cov.boot,w.cov.boot~fc)
  mod.lm <- summary(lm.beta(mod.lm))
  beta.val <- mod.lm$coefficients["fc","Standardized"]

  return(beta.val)
}


beta.ad.boot <- boot(data=df.ad,statistic=sv2a.cov.z.boot,R=1000)
beta.cn0.boot <- boot(data=df.cn0,statistic=sv2a.cov.z.boot,R=1000)

mod.t <- t.test(beta.ad.boot$t, beta.cn0.boot$t)
mod.t
```




#--- Analysis 2: network spreading Model (Shafiei, et al, Biol Psychiatry 2020) 
```{r}

## define functions 
get.network.spreading.beta.spin <- function(mean.sv2a.w,surrogate.maps,fc.thre){
  mapply(function(x){
    mean.w.ad.surrog.curren <- surrogate.maps[,x]
    
    fc.vals <- fc.mat[as.vector(lower.tri(fc.mat))]
    fc.thre.val <- fc.vals[order(fc.vals)][length(fc.vals)*fc.thre+1]

    fc.w.sv2a.ad <- mapply(function(k){
      fc.tmp <- fc.mat
      fc.tmp[which(fc.tmp<fc.thre.val)] <- NA
      fc.w.tmp <- mean(mean.w.ad.surrog.curren[-k]*fc.tmp[-k,k],na.rm=T)
      fc.w.tmp
    }, 1:200)
    
    df.stat.spin <- data.frame(fc.wei.w.ad= fc.w.sv2a.ad,
                               mean.w = mean.sv2a.w)
    
    mod.lm <- lm(data = df.stat.spin,mean.w~fc.wei.w.ad)
    mod.lm <- summary(lm.beta(mod.lm))
    beta.val.spin <- mod.lm$coefficients[2,2]
    
    return(beta.val.spin)
  }, 1:1000)
}

get.network.spreading.beta.shuffled <- function(mean.sv2a.w, fc.thre){
  mapply(function(x){
    fc.mat.shuffled <- as.matrix(shuffled.fc[[x]])
    
    fc.shuffled.vals <- fc.mat.shuffled[as.vector(lower.tri(fc.mat.shuffled))]
    fc.shuffled.thre.val <- fc.shuffled.vals[order(fc.shuffled.vals)][length(fc.shuffled.vals)*fc.thre+1]
    
    fc.w.sv2a.ad.shuffled <- mapply(function(k){
      fc.tmp <- fc.mat.shuffled
      fc.tmp[which(fc.tmp<fc.shuffled.thre.val)] <- NA
      fc.w.tmp <- mean(mean.sv2a.w[-k]*fc.tmp[-k,k],na.rm=T)
      fc.w.tmp
    }, 1:200)
    
    df.stat.shuffled <- data.frame(fc.wei.w.ad= fc.w.sv2a.ad.shuffled,
                                   mean.w = mean.sv2a.w)
    
    mod.lm.shuffled <- lm(data = df.stat.shuffled,mean.w~fc.wei.w.ad)
    mod.lm.shuffled <- summary(lm.beta(mod.lm.shuffled))
    beta.val.shuffled <- mod.lm.shuffled$coefficients[2,2]
 
    return(beta.val.shuffled)
  }, 1:1000)
}



## analyses 
fc.thre.all <- seq(0,1,0.25)[1:4] # set different functional network sparsity

beta.null.all <- list()
beta.empirical.all <- c()
p.spin.all <- c()
p.shuffled.all <- c()

for (i in 1:4) {
  ### empirical model
  fc.thre <- fc.thre.all[i]
  fc.vals <- fc.mat[as.vector(lower.tri(fc.mat))]
  fc.thre.val <- fc.vals[order(fc.vals)][length(fc.vals)*fc.thre+1]
  
  
  fc.w.sv2a.tmp <- mapply(function(x){
    fc.tmp <- fc.mat
    fc.tmp[which(fc.tmp<fc.thre.val)] <- NA
    fc.w.tmp <- mean(mean.w.ad[-x]*fc.tmp[-x,x],na.rm=T)
    c(fc.w.tmp)
  }, 1:200)
  
  df.stat.tmp <- data.frame(fc.wei.w = fc.w.sv2a.tmp,mean.w = mean.w.ad)
  mod.lm <- lm(data = df.stat.tmp,mean.w~fc.wei.w)
  mod.lm <- summary(lm.beta(mod.lm))
  mod.lm
  beta.val.empirical <- mod.lm$coefficients["fc.wei.w","Standardized"]
  p.val.empirical <- mod.lm$coefficients["fc.wei.w","Pr(>|t|)"]
  
  ### generate null distribution of beta values
  null.beta.spin <- get.network.spreading.beta.spin(mean.w.ad,mean.w.ad.surrogates, fc.thre)
  null.beta.shuffled <- get.network.spreading.beta.shuffled(mean.w.ad, fc.thre)
  
  df.null.beta <- data.frame(null = c(rep("spin",1000),rep("rewired",1000)),
                             beta = c(null.beta.spin,null.beta.shuffled))
  p.val.spin <- length(which(null.beta.spin > beta.val.empirical))/length(null.beta.spin)
  p.val.shuffled <- length(which(null.beta.shuffled > beta.val.empirical))/length(null.beta.shuffled)

  beta.null.all[[i]] <- df.null.beta
  beta.empirical.all[i] <- beta.val.empirical
  p.spin.all[i] <- p.val.spin
  p.shuffled.all[i] <- p.val.shuffled
}

beta.empirical.all <- data.frame(beta.empirical.all)
beta.empirical.all$x.point <- c(1,2,3,4)


df.beta.null.all <- data.frame(Sparsity=c(rep("100%",2000),rep("75%",2000),rep("50%",2000), rep("25%",2000)),
                               null.model=rep(c(rep("spin",1000),rep("shuffled",1000)),4),
                               beta.val= c(beta.null.all[[1]][,2],beta.null.all[[2]][,2],beta.null.all[[3]][,2],beta.null.all[[4]][,2]))
df.beta.null.all$Sparsity <- factor(df.beta.null.all$Sparsity, levels = c("100%","75%","50%","25%"))

y.p.val <- min(df.beta.null.all$beta.val)

significance <- c(p.spin.all[1],p.shuffled.all[1],p.spin.all[2],p.shuffled.all[2],p.spin.all[3],p.shuffled.all[3],p.spin.all[4],p.shuffled.all[4])
significance <- as.data.frame(significance)
significance$null.model <- c(rep(c("spin","shuffled"),4))
significance$Sparsity <- c(rep(c("100%","75%","50%","25%"),each = 2))
significance$y.postion <- y.p.val

ggplot(data=df.beta.null.all)+
  aes(x=Sparsity, y=beta.val)+
  geom_point(position = position_jitterdodge(0.65),shape=16, size=0.35, alpha=0.55,aes(fill = factor(null.model),color = factor(null.model)))+
  geom_boxplot(notch = F,color="black", alpha=0,aes(fill = factor(null.model),color = factor(null.model)),linewidth=0.25)+
  ylab("Regression-derived β-value")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color= " black", fill = NA),
        axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=10),
        axis.title.y = element_text(size=14))+
  scale_color_manual(values=c("#79AF97FF", "#6A6599FF"))+
  geom_point(data=beta.empirical.all,aes(x=x.point,y=beta.empirical.all),colour="#F56F5C",size=2)+
  geom_text(data=significance,aes(label = ifelse(significance<0.001, "***",ifelse(significance<0.01,"**",ifelse(significance<0.05,"*",""))), 
                                  group = null.model,y = y.postion-0.05), 
            position = position_dodge(width = .75), vjust = 0.5,size = 16 / .pt)





## subject-level analyses
for (i in 1:4) {
  fc.thre <- fc.thre.all[i]
  print(paste0("The functional network sparsity at ", fc.thre))
  fc.vals <- fc.mat[as.vector(lower.tri(fc.mat))]
  fc.thre.val <- fc.vals[order(fc.vals)][length(fc.vals)*fc.thre+1]
  
  beta.network.spreading.sub <- ddply(df.ad,.(SubID),function(x){
    w.ad.sub <- x[,paste0("W.SDM8.V",1:200)]
    w.ad.sub <- as.vector(as.matrix(w.ad.sub))
    
    fc.w.sv2a.sub <- mapply(function(x){
      fc.tmp <- fc.mat
      fc.tmp[which(fc.tmp<fc.thre.val)] <- NA
      fc.w.tmp <- mean(w.ad.sub[-x]*fc.tmp[-x,x],na.rm=T)
      fc.w.tmp
    }, 1:200)

    df.stat.ad.sub <- data.frame(fc.wei.w.sub = fc.w.sv2a.sub,mean.w = w.ad.sub)
    
    mod.lm <- lm(data = df.stat.ad.sub,mean.w~fc.wei.w.sub)
    mod.lm <- summary(lm.beta(mod.lm))
    beta.val.sub <- mod.lm$coefficients[2,2]
    beta.val.sub
  })
  
  beta.network.spreading.sub$variable <- "beta"
  
  t.mod <- t.test(beta.network.spreading.sub$V1)
  print(t.mod)
}


```


#--- Analysis 3: epicenter connectivity-based sv2a propagation in pooled abeta+ subjects
```{r}
## define functions
get.epicenter.by.rank <- function(mean.sv2a.w){
  rank.mean.w <- as.vector(rank(mean.sv2a.w))
  fc.wei.w <- mapply(function(k){
    fc.w.tmp <- mean(mean.sv2a.w[-k]*fc.mat[-k,k])
    fc.w.tmp
  }, 1:200)
  rank.fc.wei.w <- rank(fc.wei.w)
  mean.rank.current <- (rank.mean.w+rank.fc.wei.w)/2
  
  ### generated null distribution of mean ranks
  surrogate.map <- nmnulls$vasa(mean.sv2a.w,atlas='fsLR',density='32k',parcellation=parc,seed=as.integer(45))
  
  mean.rank.null <- mapply(function(j){
    mean.w.surrog <- surrogate.map[,j]
    rank.mean.w.surrog <- as.vector(rank(mean.w.surrog))
    
    fc.wei.w.surrog<- mapply(function(k){
      fc.w.tmp <- mean(mean.w.surrog[-k]*fc.mat[-k,k])
      fc.w.tmp
    }, 1:200)
    rank.fc.wei.w.surrog <- rank(fc.wei.w.surrog)
    mean.rank <- (rank.mean.w.surrog+rank.fc.wei.w.surrog)/2
    
  }, 1:1000)
  
  rank.p.vals <- mapply(function(roi){
    p.roi <- length(which(mean.rank.current[roi] > mean.rank.null[roi,]))/length(mean.rank.null[roi,])
  }, 1:200)
  
  epicenter.roi <- rep(0,number_of_ROIs)
  epicenter.roi[which(rank.p.vals<0.05)] <- 1
  
  epicenter.roi
}


get.epi.pred.beta.shuffled <- function(epicenter.rois, mean.sv2a.w){
  mapply(function(j){
    fc.mat.shuffled <- shuffled.fc[[j]]
    
    epi.fc.mat.shuffled <- as.matrix(fc.mat.shuffled[epicenter.rois,])
    epi.fc.mat.shuffled[which(epi.fc.mat.shuffled==Inf)] <- NA
    mean.epi.fc.mat.shuffled <- colMeans(epi.fc.mat.shuffled,na.rm = T)
    
    df.test.ad.shuffled <- data.frame(epi.fc = mean.epi.fc.mat.shuffled,
                                      sv2a.wscore=mean.sv2a.w)
    mod.lm.shuffled <- lm(data = df.test.ad.shuffled,sv2a.wscore~epi.fc)
    mod.lm.shuffled <- summary(lm.beta(mod.lm.shuffled))
    beta.val <- mod.lm.shuffled$coefficients[2,2]
    beta.val
  }, 1:1000)

}



## epicenter connectivity-based prediciton of group-level sv2a w score in pooled abeta+ subjects
### get group-level epicenter ROIs
epicenter.ad <- get.epicenter.by.rank(mean.w.ad)
epicenter.ad <- which(epicenter.ad==1)

## epicenter-based prediction
epi.fc.mat.ad <- fc.mat[epicenter.ad,]
mean.epi.fc.ad <- colMeans(epi.fc.mat.ad,na.rm = T)

## prediction
beta.shuffled.ad <- get.epi.pred.beta.shuffled(epicenter.ad,mean.w.ad)

df.stat.epicenter.fc.prediction <- data.frame(epicenter.fc = mean.epi.fc.ad, mean.w = mean.w.ad)
mod.lm <- lm(data = df.stat.epicenter.fc.prediction, mean.w~epicenter.fc)
mod.lm <- summary(lm.beta(mod.lm))
mod.lm
beta.val <- mod.lm$coefficients["epicenter.fc","Standardized"]

p.val.shuffled <- length(which(beta.shuffled.ad < beta.val))/length(beta.shuffled.ad)

x.range <- range(df.stat.epicenter.fc.prediction$epicenter.fc)
y.range <- range(df.stat.epicenter.fc.prediction$mean.w)
tibble(Predicted = df.stat.epicenter.fc.prediction$epicenter.fc, Response = df.stat.epicenter.fc.prediction$mean.w, nets = as.factor(sf.assign$V1)) %>%
  ggplot(aes(x=Predicted, y = Response))+
  geom_point(shape=21, size=2,aes(fill=factor(nets)))+
  geom_smooth(method = "lm",color = "grey55")+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color= " black", fill = NA),
        aspect.ratio = 1)+
  xlab("Functional connectivity to epicenter")+
  ylab("SV2A-PET W-score")+
  scale_fill_manual(values=colors.Schaefer.networks,labels=c("Visual","Motor","DAN","VAN","Limbic","FPCN","DMN"),name="")+
  scale_y_continuous(labels = scales::number_format(accuracy = 0.01))+
  annotate("text", x =  min(x.range)+2/5*diff(x.range), y = max(y.range)+1/8*diff(y.range),
           label = paste0(paste0("italic(β)==", round(beta.val,3)), "*\',\'",
                          "~", expression(italic(p)[italic(rewired)]),ifelse(p.val.shuffled>0.001,paste0("==",round(p.val.shuffled,3)),"<0.001")),
           color = "black",parse=T, size=5)

```

```{r}
## epicenter connectivity-based prediciton of subject-level sv2a w score in pooled abeta+ subjects
### define subject-specific epicenter
 df.epicenters.subject <- ddply(df.ad,.(SubID), function(x){
      mean.w.sub <- as.vector(as.matrix(x[,paste0("W.SDM8.V",1:200)]))
      epicenters.sub <- get.epicenter.by.rank(mean.w.sub)
      epicenters.sub
    })
colnames(df.epicenters.subject)[2:201] <- paste0("Epicenter.bin.V",1:200)
df.ad.epicenter <- merge(df.ad,df.epicenters.subject, by ="SubID")

### subject-level prediction
df.stat.sub <- ddply(df.ad.epicenter,.(SubID), function(x){

  mean.w.sub <- as.vector(as.matrix(x[,paste0("W.SDM8.V",1:200)]))
  epicenters.sub <- which(x[,paste0("Epicenter.bin.V",1:200)]==1)
  
  epi.fc.mat.sub <- fc.mat[epicenters.sub,]
  mean.epi.fc.sub <- colMeans(epi.fc.mat.sub,na.rm = T)
  
  fc.quantile.thre <-  quantile(mean.epi.fc.sub)
  q1.rois <- which(mean.epi.fc.sub<fc.quantile.thre[2])
  q2.rois <- which(mean.epi.fc.sub<fc.quantile.thre[3]&mean.epi.fc.sub>=fc.quantile.thre[2])
  q3.rois <- which(mean.epi.fc.sub<fc.quantile.thre[4]&mean.epi.fc.sub>=fc.quantile.thre[3])
  q4.rois <- which(mean.epi.fc.sub>=fc.quantile.thre[4])
  
  fc.q1 <- mean(mean.epi.fc.sub[q1.rois])
  fc.q2 <- mean(mean.epi.fc.sub[q2.rois])
  fc.q3 <- mean(mean.epi.fc.sub[q3.rois])
  fc.q4 <- mean(mean.epi.fc.sub[q4.rois])

  w.q1 <- mean(mean.w.sub[q1.rois])
  w.q2 <- mean(mean.w.sub[q2.rois])
  w.q3 <- mean(mean.w.sub[q3.rois])
  w.q4 <- mean(mean.w.sub[q4.rois])
  
  epi.eu.mat.sub <- eu.d[epicenters.sub,]
  mean.epi.eu.d.sub <- colMeans(epi.eu.mat.sub,na.rm = T)
  eu.q1 <- mean(mean.epi.eu.d.sub[q1.rois])
  eu.q2 <- mean(mean.epi.eu.d.sub[q2.rois])
  eu.q3 <- mean(mean.epi.eu.d.sub[q3.rois])
  eu.q4 <- mean(mean.epi.eu.d.sub[q4.rois])
  
  df.fc.sub <- data.frame(Quantile = c("Q1","Q2","Q3","Q4"),
                          Quantile.num = 1:4,
                          FC.quantile = c(fc.q1,fc.q2,fc.q3,fc.q4),
                          mean.sv2a = c(w.q1,w.q2,w.q3,w.q4),
                          EU.dist = c(eu.q1,eu.q2,eu.q3,eu.q4))
  
})

df.stat.sub <- merge(df.ad.epicenter,df.stat.sub, by ="SubID")
mod.lmer <- lmer(data = df.stat.sub,mean.sv2a ~ FC.quantile + SEX + Age.SDM8 + DX.AB +EU.dist +(1|SubID)) ;summary(mod.lmer)


```


#--- Analysis 4: epicenter connectivity-based sv2a propagation in subtypes of abeta+ subjects
```{r}
## hierachical clustering to determine three sv2a w-score subtypes
df.cluster.ad <- df.ad.epicenter
df.cluster.ad.w <- df.cluster.ad[,paste0("W.SDM8.V",1:200)]
df.cluster.ad.w <- scale(t(df.cluster.ad.w))
df.cluster.ad.w <- t(df.cluster.ad.w)

d <- dist(df.cluster.ad.w,method = 'euclidean')
fit.hc <- hclust(d, method = "ward.D")
clusters <- cutree(fit.hc,k=3)
df.ad.epicenter$clusters <- clusters


## group-level epicenter connectivity-based prediction of sv2a w-score for each subtypese
for (i in 1:3) {
  print(paste0("Analysis for cluster ", i))
  df.ad.epicenter.cluster <- subset(df.ad.epicenter, clusters==i)
  mean.w.ad.cluster <- colMeans(df.ad.epicenter.cluster[,paste0("W.SDM8.V",1:200)])
  epicenter.ad.cluster <- get.epicenter.by.rank(mean.w.ad.cluster)
  epicenter.ad.cluster <- which(epicenter.ad.cluster==1)
  
  ### epicenter-based prediction
  epi.fc.mat.ad.cluster <- fc.mat[epicenter.ad.cluster,]
  mean.epi.fc.ad.cluster <- colMeans(epi.fc.mat.ad.cluster,na.rm = T)
  
  ### prediction
  beta.shuffled.ad.cluster <- get.epi.pred.beta.shuffled(epicenter.ad.cluster, mean.w.ad.cluster)
  
  df.stat.cluster.epicenter.fc.prediction <- data.frame(epicenter.fc = mean.epi.fc.ad.cluster, mean.w = mean.w.ad.cluster)
  mod.lm <- lm(data = df.stat.cluster.epicenter.fc.prediction, mean.w~epicenter.fc)
  mod.lm <- summary(lm.beta(mod.lm))
  print(mod.lm)
  beta.val <- mod.lm$coefficients["epicenter.fc","Standardized"]
  p.val.shuffled <- length(which(beta.shuffled.ad.cluster < beta.val))/length(beta.shuffled.ad.cluster)
  
  
  ### sliding window analyses
  df.ad.epicenter.cluster$mean.sv2a <- rowMeans(df.ad.epicenter.cluster[,paste0("W.SDM8.V",1:200)])
  sub.order.cluster <- order(df.ad.epicenter.cluster$mean.sv2a,decreasing = T)
  df.ad.epicenter.cluster.slide <- df.ad.epicenter.cluster[sub.order.cluster,]
  num.sub <- length(sub.order.cluster)
  win.wid <- 10 # set the width of windows at 10 subjects
  win.num <- 9  # set the number of windows at 10 windows
  sep <- win.wid-((win.wid*win.num-num.sub)/(win.num-1))
  windows.assign <- rollapply(1:num.sub, win.wid, by = sep, c)

  
  for (i.win in 1:win.num) {
    df.current.window <- df.ad.epicenter.cluster.slide[as.vector(windows.assign[i.win,]),]
    mean.w.current.window <- colMeans(df.current.window[,paste0("W.SDM8.V",1:200)])
    epicenters.current.window <- epicenter.ad.cluster
    epi.fc.current.window <- fc.mat[epicenters.current.window,]
    
    df.test.current.window <- data.frame(epicenter.fc = colMeans(epi.fc.current.window,na.rm = T),
                                         mean.w = mean.w.current.window)
    
    beta.shuffled.current.window <- get.epi.pred.beta.shuffled(epicenters.current.window,mean.w.current.window)
    mod.lm <- lm(data = df.test.current.window,mean.w~epicenter.fc)
    mod.lm <- summary(lm.beta(mod.lm))
    beta.val <- mod.lm$coefficients["epicenter.fc","Standardized"]
    p.val.shuffled <- length(which(beta.shuffled.current.window< beta.val))/length(beta.shuffled.current.window)
    r2 <- mod.lm$adj.r.squared
    print(paste0("Sliding window analyses for cluster ", x , " window ", i.win))
    print(c(beta.val,p.val.shuffled,r2))
  }
}

```

```{r}
## subject-level epicenter connectivity-based prediction of sv2a w-score for each subtypese
for (i in 1:3) {
  print(paste0("Subject-level analysis for cluster ", i))
  df.ad.epicenter.cluster <- subset(df.ad.epicenter, clusters==i)
  
  df.stat.sub.cluster <- ddply(df.ad.epicenter.cluster,.(SubID), function(x){
    
    mean.w.sub <- as.vector(as.matrix(x[,paste0("W.SDM8.V",1:200)]))
    epicenters.sub <- which(x[,paste0("Epicenter.bin.V",1:200)]==1)
    
    epi.fc.mat.sub <- fc.mat[epicenters.sub,]
    mean.epi.fc.sub <- colMeans(epi.fc.mat.sub,na.rm = T)
    
    fc.quantile.thre <-  quantile(mean.epi.fc.sub)
    q1.rois <- which(mean.epi.fc.sub<fc.quantile.thre[2])
    q2.rois <- which(mean.epi.fc.sub<fc.quantile.thre[3]&mean.epi.fc.sub>=fc.quantile.thre[2])
    q3.rois <- which(mean.epi.fc.sub<fc.quantile.thre[4]&mean.epi.fc.sub>=fc.quantile.thre[3])
    q4.rois <- which(mean.epi.fc.sub>=fc.quantile.thre[4])
    
    fc.q1 <- mean(mean.epi.fc.sub[q1.rois])
    fc.q2 <- mean(mean.epi.fc.sub[q2.rois])
    fc.q3 <- mean(mean.epi.fc.sub[q3.rois])
    fc.q4 <- mean(mean.epi.fc.sub[q4.rois])
    
    w.q1 <- mean(mean.w.sub[q1.rois])
    w.q2 <- mean(mean.w.sub[q2.rois])
    w.q3 <- mean(mean.w.sub[q3.rois])
    w.q4 <- mean(mean.w.sub[q4.rois])
    
    
    epi.eu.mat.sub <- eu.d[epicenters.sub,]
    mean.epi.eu.d.sub <- colMeans(epi.eu.mat.sub,na.rm = T)
    eu.q1 <- mean(mean.epi.eu.d.sub[q1.rois])
    eu.q2 <- mean(mean.epi.eu.d.sub[q2.rois])
    eu.q3 <- mean(mean.epi.eu.d.sub[q3.rois])
    eu.q4 <- mean(mean.epi.eu.d.sub[q4.rois])
    
    df.fc.sub <- data.frame(Quantile = c("Q1","Q2","Q3","Q4"),
                            Quantile.num = 1:4,
                            FC.quantile = c(fc.q1,fc.q2,fc.q3,fc.q4),
                            mean.sv2a = c(w.q1,w.q2,w.q3,w.q4),
                            EU.dist = c(eu.q1,eu.q2,eu.q3,eu.q4))
    
  })
  
  df.stat.sub.cluster <- merge(df.ad.epicenter.cluster,df.stat.sub.cluster, by ="SubID")
  mod.lmer <- lmer(data = df.stat.sub.cluster,mean.sv2a ~ FC.quantile + SEX + Age.SDM8 + DX.AB +EU.dist+(1|SubID)) ;print(summary(mod.lmer))
  
}

```


#--- Analysis 5: epicenter connectivity-based prediciton of longitudinal change rates of SV2A wsocre
```{r}
## define function 
get.epi.pred.beta.followup.shuffled <- function(epicenter.rois, mean.sv2a.w, mean.sv2a.w.cr){
  mapply(function(j){
    fc.mat.shuffled <- shuffled.fc[[j]]
    
    epi.fc.mat.shuffled <- as.matrix(fc.mat.shuffled[epicenter.rois,])
    epi.fc.mat.shuffled[which(epi.fc.mat.shuffled==Inf)] <- NA
    mean.epi.fc.mat.shuffled <- colMeans(epi.fc.mat.shuffled,na.rm = T)
    
    df.test.ad.shuffled <- data.frame(epi.fc = mean.epi.fc.mat.shuffled,
                                      sv2a.wscore=mean.sv2a.w,
                                      sv2a.wscore.cr = mean.sv2a.w.cr)
    mod.lm.shuffled <- lm(data = df.test.ad.shuffled,sv2a.wscore~epi.fc+mean.sv2a.w.cr)
    mod.lm.shuffled <- summary(lm.beta(mod.lm.shuffled))
    beta.val <- mod.lm.shuffled$coefficients[2,2]
    beta.val
  }, 1:1000)

}

## group-level analyses
df.ad.followup <- subset(df.ad.epicenter, !is.na(CR.W.SDM8.V1))
mean.w.bl <- colMeans(df.ad.followup[,paste0("W.SDM8.V",1:200)])
mean.w.cr <- colMeans(df.ad.followup[,paste0("CR.W.SDM8.V",1:200)])
epicenter.bl <- get.epicenter.by.rank(mean.w.bl)
epicenter.bl <- which(epicenter.bl==1)

epi.fc.mat.bl <- fc.mat[epicenter.bl,]
mean.epi.fc.bl <- colMeans(epi.fc.mat.bl,na.rm = T)

beta.shuffled.followup <- get.epi.pred.beta.followup.shuffled(epicenter.bl,mean.w.bl,mean.w.cr)

df.test.followup <- data.frame(epicenter.fc = mean.epi.fc.bl,
                               sv2a.wscore.cr = mean.w.cr,
                               sv2a.wscore.bl = mean.w.bl)

mod.lm <- lm(data = df.test.followup,sv2a.wscore.cr~epicenter.fc+sv2a.wscore.bl)
mod.lm <- summary(lm.beta(mod.lm));print(mod.lm)
beta.val <- mod.lm$coefficients["epicenter.fc","Standardized"]
p.val.shuffled <- length(which(beta.shuffled.followup< beta.val))/length(beta.shuffled.followup)

```

```{r}
## subject-level analyses
beta.sub.followup <- ddply(df.ad.followup,.(SubID), function(x){
  mean.w.sub.cr <- as.vector(as.matrix(x[,paste0("CR.W.SDM8.V",1:200)]))
  mean.w.sub.bl <- as.vector(as.matrix(x[,paste0("W.SDM8.V",1:200)]))
  epicenters.bl.sub <- which(x[,paste0("Epicenter.bin.V",1:200)]==1)
  
  epi.fc.mat.sub <- fc.mat[epicenters.bl.sub,]
  mean.epi.fc.sub <- colMeans(epi.fc.mat.sub,na.rm = T)
  
  df.sub.tmp <- data.frame(mean.w.sub.cr = mean.w.sub.cr,
                           mean.w.sub.bl = mean.w.sub.bl,
                           mean.fc.sub = mean.epi.fc.sub)
  beta.sub <- coefficients(lm.beta(lm(data=df.sub.tmp,mean.w.sub.cr~mean.fc.sub+mean.w.sub.bl)))[2]
  c(beta.sub)
})

beta.sub.followup$variable <- "beta"
t.mod <- t.test(beta.sub.followup$mean.fc.sub)
t.mod

```



#--- Analysis 6: effect of plasma ptau concentration on connectivity-mediated synaptic loss propagation
```{r}
## determine subject-specific epicenters for all participants 
df.epicenters.subject.all <- ddply(df,.(SubID), function(x){
      mean.w.sub <- as.vector(as.matrix(x[,paste0("W.SDM8.V",1:200)]))
      epicenters.sub <- get.epicenter.by.rank(mean.w.sub)
      epicenters.sub
    })
colnames(df.epicenters.subject.all)[2:201] <- paste0("Epicenter.bin.V",1:200)
df.epicenter <- merge(df,df.epicenters.subject.all, by ="SubID")

## derive subject-level connectivity-based synaptic loss
beta.sub.all <- ddply(df.epicenter,.(SubID), function(x){
  mean.w.sub <- as.vector(as.matrix(x[,paste0("W.SDM8.V",1:200)]))
  epicenters.sub <- which(x[,paste0("Epicenter.bin.V",1:200)]==1)
  
  epi.fc.mat.sub <- fc.mat[epicenters.sub,]
  mean.epi.fc.sub <- colMeans(epi.fc.mat.sub,na.rm = T)
  
  df.sub.tmp <- data.frame(mean.w.sub = mean.w.sub,
                           mean.fc.sub = mean.epi.fc.sub)
  beta.sub <- coefficients(lm.beta(lm(data=df.sub.tmp,mean.w.sub~mean.fc.sub)))[2]
  c(beta.sub)
})
colnames(beta.sub.all)[2] <- "connectivity.SV2A.beta"
t.test(beta.sub.all$connectivity.SV2A.beta)

df.all.ptau181 <- merge(df.epicenter, beta.sub.all, by ="SubID")
df.ad.ptau181  <- subset(df.all.ptau181, AV45.bin==1)


## linear regression
mod.all <- lm(data = df.all.ptau181,connectivity.SV2A.beta~ log(pTau181)+ Age.SDM8 +AV45.global.mean+ SEX + DX); summary(lm.beta(mod.all))
mod.ad <- lm(data = df.ad.ptau181,connectivity.SV2A.beta~ log(pTau181)+ Age.SDM8 +AV45.global.mean+ SEX + DX); summary(lm.beta(mod.ad))

```

```{r}
## interaction effect between ptau181 and epicenter functional connectivity
### pooled sample
df.stat.plasma.ptau181.interaction <- subset(df.stat.sub, !is.na(pTau181))
df.stat.plasma.ptau181.interaction$pTau181.bin <- ifelse(df.stat.plasma.ptau181.interaction$pTau181>= median(df.stat.plasma.ptau181.interaction$pTau181, na.rm = T),1,0)

mod.lmer <- lmer(data = df.stat.plasma.ptau181.interaction,
                 mean.sv2a ~ FC.quantile*log(pTau181) + SEX + Age.SDM8 + DX + AV45.global.mean + EU.dist +(1|SubID))
summary(mod.lmer)

y.range <- range(df.stat.plasma.ptau181.interaction$mean.sv2a)

ggplot(data=df.stat.plasma.ptau181.interaction)+
  aes(x=Quantile ,y=mean.sv2a, color=factor(pTau181.bin))+
  geom_point(shape=19,alpha=0.5)+
  geom_line(alpha=0.15,aes(group=SubID))+
  geom_smooth(method = "lm",aes(x=Quantile.num , y = mean.sv2a, color =factor(pTau181.bin)), linewidth=0.5)+
  labs(x = "",y = "SV2A-PET W-score") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color= " black", fill = NA),
        axis.text.x = element_text(size=12))+
  scale_x_discrete()+
  scale_color_nejm(name="Plasma p-tau181",labels=c("<median",">median"))+
  guides(fill = "none")

### abeta positive sample
df.stat.plasma.ptau181.interaction.ad <- subset(df.stat.sub, !is.na(pTau181)& AV45.bin==1)
df.stat.plasma.ptau181.interaction.ad$pTau181.bin <- ifelse(df.stat.plasma.ptau181.interaction.ad$pTau181>= median(df.stat.plasma.ptau181.interaction.ad$pTau181, na.rm = T),1,0)

mod.lmer <- lmer(data = df.stat.plasma.ptau181.interaction.ad,
                 mean.sv2a ~ FC.quantile*log(pTau181) + SEX + Age.SDM8 + DX +AV45.global.mean + EU.dist +(1|SubID))
summary(mod.lmer)

y.range <- range(df.stat.plasma.ptau181.interaction.ad$mean.sv2a)

ggplot(data=df.stat.plasma.ptau181.interaction.ad)+
  aes(x=Quantile ,y=mean.sv2a, color=factor(pTau181.bin))+
  geom_point(shape=19,alpha=0.5)+
  geom_line(alpha=0.15,aes(group=SubID))+
  geom_smooth(method = "lm",aes(x=Quantile.num , y = mean.sv2a, color =factor(pTau181.bin)), linewidth=0.5)+
  labs(x = "",y = "SV2A-PET W-score") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        panel.border = element_rect(color= " black", fill = NA),
        axis.text.x = element_text(size=12))+
  scale_x_discrete()+
  scale_color_nejm(name="Plasma p-tau181",labels=c("<median",">median"))+
  guides(fill = "none")
```